---
title: "PACCs analysis of scRNA PC3/MDA cells w/ Cisplatin/Docetaxel and validation in scRNA and bulk RNA for mCRPC"
output: html_notebook
---
#Load libraries
```{r}

library(Seurat)
library(copykat)
library(tidyverse)
library(SeuratDisk)
library(mclust)
library(matrixStats)
library(corrplot)
library(biomaRt)
library(ggpubr)
library(msigdbr)
library(ggsci)
library(GSVA)
library(GSEABase)
library(data.table)
options(future.globals.maxSize = 4000 * 1024^5)
library(MAST)
library(rcartocolor)
hallmark_sets <- msigdbr(species = "Homo sapiens", category = "H")
#hallmark_sets %>% saveRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/data/hallmark_pathways.RDS")

```


# ==============================================================================================================================
# Section 1:Processing copyKAT results for Docetaxel/Cisplatin treated PC3 cells, plotting and creating features for classifiers
# ==============================================================================================================================
#Select colors
```{r}

# Extract AAAS palette
aaas_colors <- pal_aaas("default")(10)  # 10 predefined AAAS colors
# Extend it by repeating or rotating
n_needed <- 100
my_colors <- rep(aaas_colors, length.out = n_needed)


```


#Load Hallmark pathways
```{r}
hallmark_sets <- msigdbr(species = "Homo sapiens", category = "H")
#Create gene sets associated with PACCs
#paccs_sets <- filter(hallmark_sets, grepl(pattern = "STAT5|INTERFERON|COAGULATION|APOPTOSIS|AUTOPHAGY|INFLAMMATORY|TNF|COMPLEMENT", x = gs_name)) %>%
 # group_by(gs_name) %>%
#  summarise(genes = list(unique(gene_symbol))) %>%
  #deframe()
paccs_sets <- hallmark_sets %>%
  group_by(gs_name) %>%
  summarise(genes = list(unique(gene_symbol))) %>%
  deframe()

```


#Feature calculation for Docetaxel treated PC3 cells
```{r}
#Download downsampled seurat object
scrna <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/pc3_analysis//scrna_ds_pc3_d.rds")
scrna <- subset(scrna, subset = Notation %in% c("PC3", "d1", "d5", "d10"))
scrna <- scrna %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData()


#Module scoring stored in metadata
scrna <- AddModuleScore(object = scrna, features = paccs_sets, nbin = 10, name = names(paccs_sets))

#Load copykat results
copykat.res <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/pc3_analysis/copykatres_pc3_d.rds")
sc.meta <- scrna@meta.data

# Extract just the numeric CNV calls (columns = cells)
chrom_info <- copykat.res[[1]]$CNAmat[, 1]   # Assuming first column is chromosome names
cnv_numeric <- copykat.res[[1]]$CNAmat[, -(1:3)] #Columns as cells rows as copynumber calls relative to reference for each bin

# Sum of absolute CNV deviation per cell
cnv_burden <- colSums(abs(as.matrix(cnv_numeric)), na.rm = TRUE)

# Convert to data.table for plotting
cnv_summary <- data.frame(
  cell = names(cnv_burden),
  abs_cnv = cnv_burden
)

#Add gain and losses 
cnv_gains <- colSums(cnv_numeric > 0.1, na.rm = TRUE)
cnv_losses <- colSums(cnv_numeric < -0.1, na.rm = TRUE)
cnv_summary$gains <- cnv_gains[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary <- mutate(cnv_summary, cnv_total = gains + losses)

#Merge metadata from seurat object w/ CNV calls - left joined so if there is no cnv data then just NA for that cell and dropped in downstream analysis
sc.meta <- sc.meta %>% mutate(cell = sub(pattern = "-", replacement = "\\.", x = row.names(.))) %>% left_join(cnv_summary)

#Conver to long format
sc.meta.long <- sc.meta  %>% dplyr::select(Percent.MT, S.Score, cnv_total, nCount_RNA,nFeature_RNA, G2M.Score, cell, Notation, starts_with("HALLMARK")) %>% pivot_longer(cols = -c("cell", "Notation")) %>% mutate(Notation = ifelse(is.na(Notation), "Spike-in", Notation)) %>% mutate(Notation = factor(x = Notation, levels = c("Spike-in", "PC3", "d1", "d5", "d10")))


#Plot features across treatments
plt <- sc.meta.long %>% 
  # 1. Filter out Hallmark genes
  dplyr::filter(!grepl(pattern = "HALLMARK", x = name)) %>%
  # 2. Rename 'name' values for the facet and 'Notation' for the axis
  mutate(
    name = ifelse(name == "cnv_total", "Total CNV", name), # Replace 'old_name_here'
    Notation = recode(Notation, 
                      "PC3" = "Untreated", 
                      "d1" = "D1", 
                      "d5" = "D5", 
                      "d10" = "D10"),
    Notation = factor(Notation, levels = c("Untreated", "D1", "D5", "D10"))
  ) %>%
  ggplot(aes(x = Notation, y = value, fill = name)) +
  geom_boxplot(outlier.shape = NA) +
  theme_bw(base_size = 20) +
  ylab("Features") +
  scale_fill_viridis_d(option = "D") +
  facet_wrap(~ name, scale = "free") +
  theme(
    legend.position = "none", 
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
ggsave("/Users/debanjansaha/Desktop/pc3_cnv_boxplot_d.pdf", plt, width = 10)

#Write metadata to csv file - filter out Spike-in and C1 cells (unclear if these undergo WGD)
#sc.meta %>% dplyr::select(Percent.MT, S.Score, cnv_total, nCount_RNA,nFeature_RNA, G2M.Score, cell, Notation, starts_with("HALLMARK")) %>% filter(!(Notation %in% c(NA, "d1") ))%>% write.csv(file = "/Users/debanjansaha/Desktop/copykatres_pc3_d_meta.csv")
#ggsave("/Users/debanjansaha/Desktop/pc3_cnv_boxplot_d.pdf", plt)


```

#Plots for CN profile of Docetaxel treated PC3 cells
```{r}

copykat.res <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/pc3_analysis//copykatres_pc3_d.rds")

avg_cnv <- copykat.res[[1]]$CNAmat %>% mutate(coord = paste0(chrom, ".", chrompos)) %>% dplyr::select(-c(1:3)) %>% column_to_rownames(var = "coord") %>% t() %>% as.data.frame() %>% mutate(cell = row.names(.)) %>% mutate(Notation = case_when(
  grepl(pattern = "untreated_", x = cell) ~"Untreated",
  grepl(pattern = "c1_", x = cell) ~"C1",
  grepl(pattern = "c5_", x = cell) ~ "C5",
  grepl(pattern = "c10_", x = cell) ~ "C10",
  grepl(pattern = "d1_", x = cell) ~ "D1",
  grepl(pattern = "d5_", x = cell) ~ "D5",
  grepl(pattern = "d10_", x = cell) ~ "D10",
  TRUE ~ "Spike-in"
  
)) %>%
  dplyr::select(-cell) %>%                    
  group_by(Notation) %>%
  summarize(across(everything(), mean, na.rm = TRUE)) 

# Ensure chrom is character BEFORE factorizing
avg_cnv <- avg_cnv %>%
  pivot_longer(cols = -Notation) %>%
  separate(name, into = c("chrom", "pos"), sep = "\\.", convert = FALSE) %>%  # <- DO NOT convert
  mutate(
    chrom = factor(chrom, levels = as.character(1:23)),  # correct levels
    pos = as.integer(pos)
  ) %>%
  arrange(chrom, pos) %>%
  mutate(genome_x = dense_rank(interaction(chrom, pos, lex.order = TRUE)))

# Compute chromosome label positions
chrom_labels <- avg_cnv %>%
  group_by(chrom) %>%
  summarize(midpoint = median(genome_x)) 


# Plot with chromosome labels on x-axis
plt <- avg_cnv %>%
  ggplot(aes(x = genome_x, y = value, color = Notation)) +
  ggpubr::theme_pubr(base_size = 20) +
  scale_x_continuous(
    name = "Chromosome",
    breaks = chrom_labels$midpoint,
    labels = chrom_labels$chrom
  ) +
  labs(y = "CN Deviation") + theme(axis.text.x = element_text(angle = 90)) + geom_line(alpha = 0.8, size = 0.8) + rcartocolor::scale_color_carto_d(palette = "Bold")
plt
ggsave("/Users/debanjansaha/Desktop/pc_plots_v4/pc3_cnplot_d.pdf", plt, width = 15)


chrom = "2"
plt <- avg_cnv %>% filter(chrom == 2) %>% ggplot(aes(x = order(pos, genome_x), y = value, color = Notation)) + ggpubr::theme_pubr(base_size = 20) + geom_line(alpha = 0.8, size = 0.8) + xlab(paste0("Chromosome: ", chrom)) + ylab("CN deviation")+ rcartocolor::scale_color_carto_d(palette = "Bold")
ggsave(paste0("/Users/debanjansaha/Desktop/pc_plots_v4/pc3_cnplot_d_", chrom, ".pdf"), plt, width = 10)



```


#Feature calculation for Cisplatin treated PC3 cells
```{r}
#Read in downsampled seurat object
scrna <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/pc3_analysis//scrna_ds_pc3_c.rds")
#hallmark_sets <- msigdbr(species = "Homo sapiens", category = "H")
scrna <- subset(scrna, subset = Notation %in% c("PC3", "c1", "c5", "c10"))
scrna <- scrna %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData()
scrna <- AddModuleScore(object = scrna, features = paccs_sets, name = names(paccs_sets), nbin = 10)


#Run PCA
scrna <- RunPCA(scrna)

#Run UMAP
scrna <- RunUMAP(scrna, dims = 1:20, return.model = T)

#Copy UMAP embeddings
umap_coords <- Embeddings(scrna, "umap")
scrna$UMAP_1 <- umap_coords[, 1]
scrna$UMAP_2 <- umap_coords[, 2]

#Load copykat results
copykat.res <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/pc3_analysis//copykatres_pc3_c.rds")
sc.meta <- scrna@meta.data

#Extract just the numeric CNV calls (columns = cells)
chrom_info <- copykat.res[[1]]$CNAmat[, 1]   # Assuming first column is chromosome names
cnv_numeric <- copykat.res[[1]]$CNAmat[, -(1:3)] #Columns as cells rows as copynumber calls relative to reference for each bin

#Sum of absolute CNV deviation per cell
cnv_burden <- colSums(abs(as.matrix(cnv_numeric)), na.rm = TRUE)

#Convert to data.table for plotting
cnv_summary <- data.frame(
  cell = names(cnv_burden),
  abs_cnv = cnv_burden
)

#Add gain and losses 
cnv_gains <- colSums(cnv_numeric > 0.1, na.rm = TRUE)
cnv_losses <- colSums(cnv_numeric < -0.1, na.rm = TRUE)
cnv_summary$gains <- cnv_gains[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary <- mutate(cnv_summary, cnv_total = gains + losses)

#Merge metadata from seurat object w/ CNV calls - left joined so if there is no cnv data then just NA for that cell and dropped in downstream analysis
sc.meta <- sc.meta %>% mutate(cell = sub(pattern = "-", replacement = "\\.", x = row.names(.))) %>% left_join(cnv_summary)

sc.meta.long <- sc.meta  %>% 
  dplyr::select(Percent.MT, S.Score, cnv_total, nCount_RNA,nFeature_RNA, G2M.Score, cell, Notation, starts_with("HALLMARK")) %>% pivot_longer(cols = -c("cell", "Notation")) %>% mutate(Notation = ifelse(is.na(Notation), "Spike-in", Notation)) %>% 
  mutate(Notation = factor(x = Notation, levels = c("PC3", "c1", "c5", "c10")))  

#Plot features across treatments
 
plt <-  sc.meta.long %>% 
  # 1. Filter out Hallmark genes
  dplyr::filter(!grepl(pattern = "HALLMARK", x = name)) %>%
  # 2. Rename 'name' values for the facet and 'Notation' for the axis
  mutate(
    name = ifelse(name == "cnv_total", "Total CNV", name), # Replace 'old_name_here'
    Notation = recode(Notation, 
                      "PC3" = "Untreated", 
                      "c1" = "C1", 
                      "c5" = "C5", 
                      "c10" = "C10"),
    Notation = factor(Notation, levels = c("Untreated", "C1", "C5", "C10"))
  ) %>%
  ggplot(aes(x = Notation, y = value, fill = name)) +
  geom_boxplot(outlier.shape = NA) +
  theme_bw(base_size = 20) +
  ylab("Features") +
  scale_fill_viridis_d(option = "D") +
  facet_wrap(~ name, scale = "free") +
  theme(
    legend.position = "none", 
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

#Write metadata to csv file - filter out Spike-in and C1 cells (unclear if these undergo WGD)
#sc.meta %>% dplyr::select(c(Percent.MT, S.Score, cnv_total, nCount_RNA,nFeature_RNA, G2M.Score, cell, Notation, starts_with("HALLMARK"))) %>% filter(!(Notation %in% c(NA, "c1") ))%>% write.csv(file = "/Users/debanjansaha/Desktop/copykatres_pc3_c_meta.csv")


ggsave("/Users/debanjansaha/Desktop/pc_plots_v4/pc3_cnv_boxplot_c.pdf", plt, width = 10)



```



#Plots for CN profile of Cisplatin treated PC3 cells
```{r}
copykat.res <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/pc3_analysis//copykatres_pc3_c.rds")

avg_cnv <- copykat.res[[1]]$CNAmat %>% mutate(coord = paste0(chrom, ".", chrompos)) %>% dplyr::select(-c(1:3)) %>% column_to_rownames(var = "coord") %>% t() %>% as.data.frame() %>% mutate(cell = row.names(.)) %>% mutate(Notation = case_when(
  grepl(pattern = "untreated_", x = cell) ~"Untreated",
  grepl(pattern = "c1_", x = cell) ~"Cisplatin 1DPT",
  grepl(pattern = "c5_", x = cell) ~ "Cisplatin 5DPT",
  grepl(pattern = "c10_", x = cell) ~ "Cisplatin 10DPT",
  grepl(pattern = "d1_", x = cell) ~ "Docetaxel 1DPT",
  grepl(pattern = "d5_", x = cell) ~ "Docetaxel 5DPT",
  grepl(pattern = "d10_", x = cell) ~ "Docetaxel 10",
  TRUE ~ "Spike-in"
  
)) %>%
  dplyr::select(-cell) %>%                    
  group_by(Notation) %>%
  summarize(across(everything(), mean, na.rm = TRUE)) 

# Ensure chrom is character BEFORE factorizing
avg_cnv <- avg_cnv %>%
  pivot_longer(cols = -Notation) %>%
  separate(name, into = c("chrom", "pos"), sep = "\\.", convert = FALSE) %>%  # <- DO NOT convert
  mutate(
    chrom = factor(chrom, levels = as.character(1:23)),  # correct levels
    pos = as.integer(pos)
  ) %>%
  arrange(chrom, pos) %>%
  mutate(genome_x = dense_rank(interaction(chrom, pos, lex.order = TRUE)))

# Compute chromosome label positions
chrom_labels <- avg_cnv %>%
  group_by(chrom) %>%
  summarize(midpoint = median(genome_x)) 
avg_cnv$Notation <- factor(avg_cnv$Notation, 
                      levels = c( "Spike-in", "Untreated",
                                 "Cisplatin 1DPT", "Cisplatin 5DPT", "Cisplatin 10DPT"))

# Plot with chromosome labels on x-axis
plt <- avg_cnv %>%
  ggplot(aes(x = genome_x, y = value, color = Notation)) +
  ggpubr::theme_pubr(base_size = 20) +
  scale_x_continuous(
    name = "Chromosome",
    breaks = chrom_labels$midpoint,
    labels = chrom_labels$chrom
  ) +
  labs(y = "CN Deviation") + theme(axis.text.x = element_text(angle = 90)) + geom_line(alpha = 0.8, size = 0.8) + scale_color_viridis_d(option = "D", direction = -1)

plt
ggsave("/Users/debanjansaha/Desktop/pc3_cnplot_c.pdf", plt, width = 15)


chrom = 2
plt <- avg_cnv %>% filter(chrom == 2) %>% ggplot(aes(x = order(pos, genome_x), y = value, color = Notation)) + theme_pubr(base_size = 20) + geom_line(alpha = 0.8, size = 0.8) + xlab(paste0("Chromosome: ", chrom)) + ylab("CN deviation") + scale_color_viridis_d(option = "D", direction = -1)
ggsave(paste0("/Users/debanjansaha/Desktop/pc_plots_v4/pc3_cnplot_c_", chrom, ".pdf"), plt, width = 11)


```



# ==============================================================================================================================
# Section 2: Processing single cell RNA data for MDA cellines and processing copyKAT output for testing classifiers
# ==============================================================================================================================

#MDA cells with cisplatin
```{r}

#Read in downsampled seurat object
scrna <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/mda_analysis/scrna_ds_mda_c.rds")
scrna <- subset(scrna, subset = Group %in% c("untreated", "c1", "c5", "c10"))
scrna <- scrna %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData()
scrna <- AddModuleScore(object = scrna, features = paccs_sets, nbin = 10, name = names(paccs_sets))


#Load copykat results
copykat.res <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/mda_analysis//copykatres_mda_c.rds")
sc.meta <- scrna@meta.data

# Extract just the numeric CNV calls (columns = cells)
chrom_info <- copykat.res[[1]]$CNAmat[, 1]   # Assuming first column is chromosome names
cnv_numeric <- copykat.res[[1]]$CNAmat[, -(1:3)] #Columns as cells rows as copynumber calls relative to reference for each bin

# Sum of absolute CNV deviation per cell
cnv_burden <- colSums(abs(as.matrix(cnv_numeric)), na.rm = TRUE)

# Convert to data.table for plotting
cnv_summary <- data.frame(
  cell = names(cnv_burden),
  abs_cnv = cnv_burden
)

#Add gain and losses 
cnv_gains <- colSums(cnv_numeric > 0.1, na.rm = TRUE)
cnv_losses <- colSums(cnv_numeric < -0.1, na.rm = TRUE)
cnv_summary$gains <- cnv_gains[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary <- mutate(cnv_summary, cnv_total = gains + losses)


#Merge metadata from seurat object w/ CNV calls - left joined so if there is no cnv data then just NA for that cell and dropped in downstream analysis
sc.meta <- sc.meta %>% mutate(cell = sub(pattern = "-", replacement = "\\.", x = row.names(.))) %>% left_join(cnv_summary)

#Save to csv
sc.meta %>% dplyr::select(Percent.MT, S.Score, cnv_total, nCount_RNA,nFeature_RNA, G2M.Score, cell, Group, starts_with("HALLMARK")) %>% filter(!(Group %in% c(NA, "c1") ))%>% write.csv(file = "/Users/debanjansaha/Desktop/copykatres_mda_c_meta.csv")
```

#MDA cells with docetaxel
```{r}

#Read in downsampled seurat object
scrna <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/mda_analysis//scrna_ds_mda_d.rds")
scrna <- subset(scrna, subset = Group %in% c("untreated", "d1", "d5", "d10"))
scrna <- scrna %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData()
scrna <- AddModuleScore(object = scrna, features = paccs_sets, nbin = 10, name = names(paccs_sets))


#Load copykat results
copykat.res <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/mda_analysis//copykatres_mda_d.rds")
sc.meta <- scrna@meta.data

# Extract just the numeric CNV calls (columns = cells)
chrom_info <- copykat.res[[1]]$CNAmat[, 1]   # Assuming first column is chromosome names
cnv_numeric <- copykat.res[[1]]$CNAmat[, -(1:3)] #Columns as cells rows as copynumber calls relative to reference for each bin

# Sum of absolute CNV deviation per cell
cnv_burden <- colSums(abs(as.matrix(cnv_numeric)), na.rm = TRUE)

# Convert to data.table for plotting
cnv_summary <- data.frame(
  cell = names(cnv_burden),
  abs_cnv = cnv_burden
)

#Add gain and losses 
cnv_gains <- colSums(cnv_numeric > 0.1, na.rm = TRUE)
cnv_losses <- colSums(cnv_numeric < -0.1, na.rm = TRUE)
cnv_summary$gains <- cnv_gains[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary$losses <- cnv_losses[cnv_summary$cell]
cnv_summary <- mutate(cnv_summary, cnv_total = gains + losses)


#Merge metadata from seurat object w/ CNV calls - left joined so if there is no cnv data then just NA for that cell and dropped in downstream analysis
sc.meta <- sc.meta %>% mutate(cell = sub(pattern = "-", replacement = "\\.", x = row.names(.))) %>% left_join(cnv_summary)


#Save to csv
sc.meta %>% dplyr::select(Percent.MT, S.Score, cnv_total, nCount_RNA,nFeature_RNA, G2M.Score, cell, Group, starts_with("HALLMARK")) %>% filter(!(Group %in% c(NA, "d1") ))%>% write.csv(file = "/Users/debanjansaha/Desktop/copykatres_mda_d_meta.csv")




```



#Sensitivity analysis


```{r}


# Combine all entries into one tidy dataframe
auc_df_d <- read.csv(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/sensitivity_analysis/auc_mda_d.csv")


auc_df_d <- auc_df_d %>%
  mutate(
    spikein   = str_extract(sample_id, "spk\\d+") %>% str_remove("spk"),
    untreated = str_extract(sample_id, "ut\\d+")  %>% str_remove("ut"),
    d1 = str_extract(sample_id, "d1\\d+")  %>% str_remove("d1"),
    d5 = str_extract(sample_id, "d5\\d+")  %>% str_remove("d5"),
    d10 = str_extract(sample_id, "d10\\d+")  %>% str_remove("d10"),  ) %>%
  rowwise() %>%
  mutate(
    Dataset_label = paste0(
      "Spikein: ", spikein,
      ", Untreated: ", untreated,
      ", 1DPT: ", d1,
      ", 5DPT: ", d5,
      ", 10DPT: ", d10)) %>%
  ungroup() %>%
  dplyr::select(-spikein, -untreated, -d1, -d5, -d10)

plt1 <- auc_df_d %>% ggplot(aes(y = Dataset_label, x = auc, color = model)) + geom_jitter() + xlim(c(0.5,1.05)) + theme_bw(base_size = 10) + scale_color_aaas(name = "Model") + ylab("MDA with Docetaxel") + xlab("AUC") + theme(legend.text = element_text(size = 12), legend.title = element_text(size = 12))


# Combine all entries into one tidy dataframe
auc_df_c <- read.csv(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/sensitivity_analysis/auc_mda_c.csv")

auc_df_c <- auc_df_c %>%
  mutate(
    spikein   = str_extract(sample_id, "spk\\d+") %>% str_remove("spk"),
    untreated = str_extract(sample_id, "ut\\d+")  %>% str_remove("ut"),
    d1 = str_extract(sample_id, "d1\\d+")  %>% str_remove("d1"),
    d5 = str_extract(sample_id, "d5\\d+")  %>% str_remove("d5"),
    d10 = str_extract(sample_id, "d10\\d+")  %>% str_remove("d10"),  ) %>%
  rowwise() %>%
  mutate(
    Dataset_label = paste0(
      "Spikein: ", spikein,
      ", Untreated: ", untreated,
      ", 1DPT: ", d1,
      ", 5DPT: ", d5,
      ", 10DPT: ", d10)) %>%
  ungroup() %>%
  dplyr::select(-spikein, -untreated, -d1, -d5, -d10)

plt2 <- auc_df_c %>% ggplot(aes(y = Dataset_label, x = auc, color = model)) + geom_jitter() + xlim(c(0.5,1.05)) + theme_bw(base_size = 10) + scale_color_viridis_d(name = "Model") + ylab("MDA with Cisplatin") + xlab("AUC") +   theme(legend.position = "none")


pdf(file = "/Users/debanjansaha/Desktop/pc_plots_v4/mda_paccs_sens.pdf", width = 10)
ggarrange(plt1, plt2, ncol = 1, nrow = 2)
dev.off()


```


# ==============================================================================================================================
# Section 3: Analysis of single cell/bulk RNA data from mCRPC data sets
# ==============================================================================================================================


```{r}
#Load data
scrna <- readRDS("/Users/debanjansaha/Desktop/mCRPC_analysis/R/crpc_analysis/results/crpc_epi20000tme20000seed123.rds")
scrna <- scrna %>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData()
scrna <- AddModuleScore(object = scrna, features = paccs_sets, nbin = 10, name = names(paccs_sets))


#Run PCA
scrna <- RunPCA(scrna)

#Run UMAP
scrna <- RunUMAP(scrna, dims = 1:20)

#Save UMAP embeddings
umap_coords <- Embeddings(scrna, "umap")
scrna$UMAP_1 <- umap_coords[, 1]
scrna$UMAP_2 <- umap_coords[, 2]
scrna.meta <- scrna@meta.data 
row.names(scrna.meta) <- scrna.meta$cell_id

#Format Metadata
#meta <- read_csv(file ="/Users/debanjansaha/Desktop/mCRPC_analysis/R/crpc_analysis/results//crpc_epi20000tme20000seed123_meta.csv" )

#Preprocess for pacc_classifier_cli.py
#pp <- dplyr::select(scrna.meta, subtype, percent.mt, S.Score, nCount_RNA,nFeature_RNA, G2M.Score,cell_id,coarse_ano, starts_with("HALLMARK"), UMAP_1, UMAP_2 ) %>%  mutate(Percent.MT = percent.mt) %>% merge(dplyr::select(meta, cell_id, cnv_total))  

#pp %>% write.csv(file ="/Users/debanjansaha/Desktop/mCRPC_analysis/R/crpc_analysis/results/crpc_epi20000tme20000seed123_pp.csv" )
scrna@meta.data
```

#Load predictions
```{r}
#Load predictions 
pred <- read_csv(file ="/Users/debanjansaha/Desktop/mCRPC_analysis/R/crpc_analysis/results//crpc_epi20000tme20000seed123_predictions.csv" )
pred <- pred %>% dplyr::select(c("cell_id", ends_with("_prob"), cnv_total)) 
row.names(pred) <- pred$cell_id

# Confirm rownames match colnames
all(rownames(pred) %in% row.names(scrna.meta))  # should be TRUE

# Step 2: Reorder metadata to match scrna@meta.data
pred <- pred[row.names(scrna.meta), , drop = FALSE]

# Step 3: Add the new metadata to the Seurat object
scrna.meta<- cbind(scrna.meta, pred)

data.frame(scrna.meta) %>%
  group_by(subtype) %>%
  summarise(
    n_missing = sum(is.na(cnv_total)),
    n_total   = n(),
    .groups = "drop"
  )

```

#Bar plots
```{r}
p1 <- data.frame(scrna.meta) %>% 
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = LogisticRegression_PACCs_prob   )) +
  geom_point(size = 0.1) + 
  labs(title = "LogReg PCs probabilities", color = "PCs Prob") + theme_bw(base_size = 20) + scale_color_viridis_c(option = "B",limits = c(0.5, 1), na.value = "grey90" ) + theme(aspect.ratio = 1)

p2 <- data.frame(scrna.meta) %>% 
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = coarse_ano)) +
  geom_point(size = 0.1) +
  labs(title = "Cell compartments", color = "Celltype") + theme_bw(base_size = 20) + scale_color_nejm() + theme(aspect.ratio = 1)

p3 <- data.frame(scrna.meta) %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = subtype)) +
  geom_point(size = 0.1)  +
  labs(title = "Histologic variant", color = "Subytpe") + theme_bw(base_size = 20) + scale_color_aaas() + theme(aspect.ratio = 1)

ggsave(p1, filename =  "/Users/debanjansaha/Desktop/pc_plots_v4/crpc_paccs_umap_1.pdf", width = 10)
ggsave(p2, filename =  "/Users/debanjansaha/Desktop/pc_plots_v4/crpc_paccs_umap_2.pdf", width = 10)
ggsave(p3, filename =  "/Users/debanjansaha/Desktop/pc_plots_v4/crpc_paccs_umap_3.pdf", width = 10)

```

#Enrichment plots
```{r}
#PC by Cell type
p4 <- scrna.meta %>%
  dplyr::select(coarse_ano, ends_with("_prob")) %>%
  pivot_longer(cols = ends_with("_prob"), names_to = "classifier", values_to = "value") %>%
  mutate(PACCs = ifelse(value >= 0.99, "PCs", "non-PCs")) %>% 
  filter(!is.na(PACCs)) %>%
  group_by(classifier, coarse_ano, PACCs) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(proportion = n / sum(n)) %>% 
  mutate(classifier = gsub(pattern = "_PACCs_prob", replacement = "", x = classifier)) %>% 
  
  ggplot(aes(x = coarse_ano, y = proportion, fill = PACCs)) +
  geom_bar(stat = "identity", position = "stack") + 
  geom_text(
    data = \(d) filter(d, PACCs == "PCs", proportion > 0.01),  
    aes(label = paste0(round(proportion, 3) * 100, "%"),
        y = 1),                                              
    vjust = -0.3,                                            
    size = 5,
    color = "black",
    fontface = "bold"
  ) +
  facet_wrap(~ classifier, scales = "free_y") +
  scale_fill_manual(values = c("non-PCs" = "grey80", "PCs" = "navy")) + 
  theme_bw(base_size = 20) +
  labs(
    x = "Cell Type",
    y = "Proportion of Cells",
    fill = "Prediction",
    title = "PC Classification by Cell Type and Model"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 15),
    panel.grid.major = element_blank()
  )

#PC by Histology 
p5 <- scrna.meta %>%
  dplyr::select(subtype, ends_with("_prob")) %>%
  pivot_longer(cols = ends_with("_prob"), names_to = "classifier", values_to = "value") %>%
  mutate(PACCs = ifelse(value >= 0.99, "PCs", "non-PCs")) %>% 
  filter(!is.na(PACCs)) %>%
  group_by(classifier, subtype, PACCs) %>%
  summarise(n = n(), .groups = "drop_last") %>%
  mutate(proportion = n / sum(n)) %>% 
  mutate(classifier = gsub(pattern = "_PACCs_prob", replacement = "", x = classifier)) %>% 
  mutate(subtype = factor(subtype, levels = c("CSPC", "CRPC", "NEPC"))) %>%
  ggplot(aes(x = subtype, y = proportion, fill = PACCs)) +
  geom_bar(stat = "identity", position = "stack") + 
  geom_text(
    data = \(d) filter(d, PACCs == "PCs", proportion > 0.01),  
    aes(label = paste0(round(proportion, 3) * 100, "%"),
        y = 1),                                              
    vjust = -0.2,                                            
    size = 5,
    color = "black",
    fontface = "bold"
  ) +
  facet_wrap(~ classifier, scales = "free_y") +
  scale_fill_manual(values = c("non-PCs" = "grey80", "PCs" = "navy")) + 
  theme_bw(base_size = 20) +
  labs(
    x = "Histology",
    y = "Proportion of Cells",
    fill = "Prediction",
    title = "PC Classification by Histologic Variant"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 15),
    panel.grid.major = element_blank()
  )

pdf(file = "/Users/debanjansaha/Desktop/pc_plots_v4/crpc_paccs_celltype.pdf", width = 10, height = 10)
ggarrange(p4, ncol = 1, nrow = 1)
dev.off()

pdf(file = "/Users/debanjansaha/Desktop/pc_plots_v4/crpc_paccs_histo.pdf", width = 10, height = 10)
ggarrange(p5, ncol = 1, nrow = 1)
dev.off()


```




#DEG
```{r}
#PACCs DEG analysis using LogReg classifier results
scrna.meta <-  data.frame(scrna.meta) %>% mutate(PACCS = case_when(
  coarse_ano != "Epi_Neuroendo" ~ "non-PACCs",
  LogisticRegression_PACCs_prob >= 0.99 ~ "PACCs",
  LogisticRegression_PACCs_prob < 0.99 ~ "non-PACCs"
)) 

#ggplot(scrna.meta, aes(x = NeuralNet_PACCs_prob)) + geom_density()

scrna@meta.data[,"PACCS"] <- scrna.meta$PACCS
#DEG analysis
paccs_deg <- FindMarkers(
  scrna,
  ident.1 = "PACCs",
  ident.2 = "non-PACCs",
  group.by = "PACCS",   min.pct = 0.1
)

#Up and downregulated DE genes
paccs_deg_up <- paccs_deg %>% arrange(-avg_log2FC) %>% dplyr::filter(p_val_adj < 1e-5  & avg_log2FC > 4  &pct.1 > 0.1 )
paccs_deg_down <- paccs_deg %>% arrange(avg_log2FC) %>% dplyr::filter(p_val_adj < 1e-5  & avg_log2FC < -4  &pct.1 > 0.1 )
```

#Volcano plot of DEG
```{r}
mutate(paccs_deg, sig = case_when(
  row.names(paccs_deg) %in% c(row.names(paccs_deg_up), row.names(paccs_deg_down)) ~ "Sig",
  TRUE ~ "NS")) %>% ggplot( aes(x = avg_log2FC, y = -log10(p_val_adj), color = sig)) +
  geom_point(alpha = 0.6) +
  scale_color_manual(values = c("Sig" = "navy", "Sig" = "blue", "NS" = "gray")) +
  geom_vline(xintercept = c(-4, 4), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(1e-5), linetype = "dashed", color = "black") +
  labs(
    title = "Volcano Plot of PACCs DEGs",
    x = "Log2 Fold Change",
    y = "-log10 Adjusted p-value",
    color = "Regulation"
  ) +
  theme_minimal()
c(row.names(paccs_deg_up), row.names(paccs_deg_down))



```

#GSVA for Hallmark pathways for PACCs
```{r}
library(GSVA)
library(viridis)

df_long <- scrna@meta.data %>%
  dplyr::select(
    PACCS,
    cell_id,
    coarse_ano,
    subtype,
    starts_with("HALLMARK")
  ) %>%
  pivot_longer(
    cols = starts_with("HALLMARK"),
    names_to = "pathway",
    values_to = "score"
  ) %>%
  filter(
    coarse_ano == "Epi_Neuroendo",
    subtype %in% c("CRPC", "NEPC"),
    !is.na(PACCS)
  ) %>%
  mutate(
    PACCS_group = ifelse(PACCS == "PACCs", "PCs", "non-PCs")
  )

## ================================
## 2. Compute mean GSVA per group
## ================================
pathway_stats <- df_long %>%
  group_by(pathway, PACCS_group) %>%
  summarize(
    mean_score = mean(score),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = PACCS_group,
    values_from = mean_score
  ) %>%
  mutate(
    mean_diff = `PCs` - `non-PCs`
  ) %>%
  arrange(desc(mean_diff))

## ================================
## 3. Join + reorder pathways
## ================================
df_plot <- df_long %>%
  left_join(
    pathway_stats %>% dplyr::select(pathway, mean_diff),
    by = "pathway"
  ) %>%
  mutate(
    pathway = fct_reorder(pathway, mean_diff)
  )

## ================================
## 4. Plot: ranked boxplots
## ================================

p <- pathway_stats %>%
  pivot_longer(
    cols = c(`PCs`, `non-PCs`),
    names_to = "group",
    values_to = "mean_score"
  ) %>%
  mutate(
    pathway = fct_reorder(pathway, mean_diff)
  ) %>%
  ggplot(aes(
    y = pathway,
    x = mean_score,
    color = group
  )) +
  geom_line(aes(group = pathway), color = "grey70") +
  geom_point(size = 3) +
  theme_bw() +
  labs(
    x = "Mean GSVA score",
    y = "Hallmark pathway",
    color = "PCs group"
  ) + scale_color_viridis_d(end = 0.5)

ggsave(filename = "/Users/debanjansaha/Desktop/pc_plots_v4/gsva.pdf", plot = p)

```



#WCDT mCRPC RNA-seq
```{r}
#Validation of PACCs DEG on WCDT samples
#load("/Users/debanjansaha/Desktop/mCRPC_analysis/data/WCDT/alternative_promoter_data_2024/environment_deeprna_build_reproduce_20230418_NCB_revision.RData")

#Load metadata from Quigly et al. 2018 paper
wcdt_meta <- read.csv(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/paccs_plots/wcdt_mcrpc_metadata.csv")

#GSVA scoring for bulk RNA-seq data
expr_mat <- read.table(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/data/WCDT/2018_04_15_matrix_rna_tpm.txt", header = T)
expr_mat <- column_to_rownames(expr_mat, var = "IDENTIFIER")
```


```{r}
#Filter PACCs DEG ensembl ids
gene_sets <- list(PACCs_score = c(row.names(paccs_deg_up) ) )
gparams <- ssgseaParam(exprData = as.matrix(log1p(expr_mat)),
                      geneSets = gene_sets)
gsva_result <- gsva(gparams)
ucsf_lcm_tpm <- gsva_result %>% 
  t() %>% data.frame() %>% 
  rownames_to_column(var= "match_id") %>% 
  mutate(match_id = gsub(pattern = "\\.", replacement = "-", x = match_id)) %>% 
  dplyr::filter(!is.na(match_id)) %>% 
  left_join(wcdt_meta,by = c("match_id" = "IDENTIFIER")) #%>% left_join(clinical_features_eu,by = c("match_id" = "sample_id"))

#Plots for aneuploidy and polyploidy 
comparisons <- list(
  c("2", "3"),
  c("2", "4")
)

p1 <- ucsf_lcm_tpm   %>% mutate(estimated.ploidy = as.factor(round(estimated.ploidy) )) %>%  ggplot(aes(y = PACCs_score, x = estimated.ploidy)) + geom_violin(fill = viridis(n = 1))  + theme_bw(base_size = 20) + stat_compare_means(
    method = "wilcox.test",
    label = "p.format" , comparisons = comparisons) + stat_summary(
    fun = median,
    fun.min = median,
    fun.max = median,
    geom = "crossbar",
    width = 0.5,
    color = "grey"
  ) + xlab("Estimated Ploidy") + ylab("PACCs score")

comparisons <- list(
  c("0.1", "0.2"),
    c("0.1", "0.3"),
    c("0.1", "0.4"),
    c("0.1", "0.5")
)

p2 <- ucsf_lcm_tpm   %>% mutate(percent.copy.altered = as.factor(round(percent.copy.altered, 1) )) %>%  ggplot(aes(y = PACCs_score, x = percent.copy.altered)) + geom_violin(fill = viridis(n = 1, direction = -1, end = 0.5))  + scale_fill_viridis_d(option = "D") + theme_bw(base_size = 20) + stat_compare_means(
    method = "wilcox.test",
    label = "p.format" , comparisons = comparisons) + stat_summary(
    fun = median,
    fun.min = median,
    fun.max = median,
    geom = "crossbar",
    width = 0.5,
    color = "grey"
  ) + xlab("Percent Copy Altered") + ylab("PACCs score")

pdf("/Users/debanjansaha/Desktop/bulk_paccs_sv.pdf")
p3 <- ucsf_lcm_tpm %>% dplyr::select(c(PACCs_score), starts_with("X"), mutation_count) %>% mutate(mutation_count = log1p(mutation_count)) %>% pivot_longer(cols = -PACCs_score) %>% filter(name %in% c("X..deletions", "X..deletions.microhom.")) %>%   ggplot(aes(x = PACCs_score, y = value  )) + facet_wrap(
  ~ name,
  scales = "free_y",
  labeller = labeller(
    name = c(
      "mutation_count" = "Mutation count",
      "X..deletions" = "Deletions",
      "X..deletions.microhom." = "Deletions (microhomology)",
      "X..duplications" = "Duplications",
      "X..insertions" = "Insertions",
      "X..inversions" = "Inversions"
    )
  )
) + geom_point(alpha = 0.5)  + theme_bw(base_size = 20) + stat_cor(method = "pearson") + geom_smooth(method = "lm" , colour = viridis(1, begin = 0.2, option = "A")) + xlab("PCs score") + ylab("Count")
dev.off() 


p4 <- ucsf_lcm_tpm %>% dplyr::select(c(PACCs_score), starts_with("X"), mutation_count) %>% mutate(mutation_count = log1p(mutation_count)) %>% pivot_longer(cols = -PACCs_score) %>% filter(!name %in% c("X..deletions", "X..deletions.microhom.")) %>%   ggplot(aes(x = PACCs_score, y = value  )) + facet_wrap(
  ~ name,
  scales = "free_y",
  labeller = labeller(
    name = c(
      "mutation_count" = "Mutation count",
      "X..deletions" = "Deletions",
      "X..deletions.microhom." = "Deletions (microhomology)",
      "X..duplications" = "Duplications",
      "X..insertions" = "Insertions",
      "X..inversions" = "Inversions"
    )
  )
) + geom_point(alpha = 0.5)  + theme_bw(base_size = 20) + stat_cor(method = "pearson") + geom_smooth(method = "lm" , colour = viridis(1, begin = 0.3, option = "A")) + xlab("PCs score") + ylab("Count")
dev.off() 


pdf("/Users/debanjansaha/Desktop/pc_plots_v4/bulk_paccs.pdf", width = 10)
ggarrange(p1, p2, ncol = 2, nrow = 1)
dev.off()


ggsave(p3, filename = "/Users/debanjansaha/Desktop/pc_plots_v4/mut1.pdf", width = 10)
ggsave(p4, filename = "/Users/debanjansaha/Desktop/pc_plots_v4/mut2.pdf", width = 10)


```



#CNV phylo tree
```{r}
library(phangorn)
library(ggtree)
library(cluster)


copykat_crpc <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/20000_copkyat_cn_matrix.rds")
set.seed(123)
# Step 1: Get one representative cell per Seurat cluster for PACCs
pacc_meta <- scrna@meta.data %>%
  filter(coarse_ano == "Epi_Neuroendo" & !is.na(PACCS)) %>% 
  mutate(cell_name = rownames(.))



# Get medoid (most central cell) from each cluster based on UMAP
medoid_cells <- pacc_meta %>%
  group_by(seurat_clusters ,patient ) %>%
  mutate(
    # Calculate distance to cluster center
    umap1_center = mean(UMAP_1),
    umap2_center = mean(UMAP_2),
    dist_to_center = sqrt((UMAP_1 - umap1_center)^2 + (UMAP_2 - umap2_center)^2)
  ) %>%
  # Pick cell closest to center
  slice_min(dist_to_center, n = 5000) %>%
  ungroup() %>%
  pull(cell_name)

print(paste("Selected", length(medoid_cells), "representative cells from", 
            length(unique(pacc_meta$seurat_clusters)), "clusters"))

# Step 2: Get CN profiles for these representative cells
cn_mat <- copykat_crpc

# Convert cell names to match copykat format
medoid_names_converted <- gsub(pattern = "-", replacement = "\\.", medoid_cells)

# Extract CN data for medoids
mediods <- cn_mat[, colnames(cn_mat) %in% medoid_names_converted]

# Remove metadata columns if present
mediods <- mediods %>%
  dplyr::select(-any_of(c("chrom", "chrompos", "abspos", "coord"))) %>%
  as.matrix()

# Remove rows with too many NAs
mediods <- mediods[rowSums(is.na(mediods)) < ncol(mediods), ]


#colnames(mediods) %>% write.table(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/crpc_analysis/results/mediod_cellnames.txt", quote = F, col.names = F, row.names = F)

```

```{r}
patients <- unique(pacc_meta$patient)
patients <- c("MSK-HP17", "MSK-HP18")

 lapply(patients, function(pid) {

  cat("Patient:", pid, "\n")

  # cells for this patient
  ids <- pacc_meta %>%
    filter(patient == pid, cell_name %in% medoid_cells) %>%
    pull(cell_name) %>%  gsub(pattern = "-", replacement = "\\.")


  # subset medoids
  m <- mediods[, colnames(mediods) %in% ids, drop = FALSE]
    if (ncol(m) < 3) {
    message("Skipping patient ", pid, " (", length(ids), " cells)")
    return(NULL)
  }
  # distance + tree
  med_dist <- dist(t(m), method = "manhattan")
  tree <- fastme.bal(med_dist)

  # annotation
  df <- pacc_meta %>%
    filter(cell_name %in% gsub(pattern = "\\.", replacement = "-", x = ids)) %>%
    mutate(label = gsub("-", "\\.", cell_name)) %>%
    filter(label %in% tree$tip.label) %>%
    dplyr::select(label, PACCS) %>%
    filter(!is.na(PACCS))

  # plot
  p <- ggtree(tree) %<+% df +
    geom_tree(aes(color = PACCS)) +
    geom_tippoint(aes(color = PACCS, alpha = PACCS), size = 2) +
    geom_treescale() +
    theme_tree2() +
    theme(legend.position = "right") +
    labs(
      color = "PCs",
      title = paste("Patient", pid)
    ) +
    scale_colour_viridis_d(option = "D", end = 0.5,breaks = c("non-PACCs", "PACCs"),                
    labels = c("Non-PC", "PC"),          
    na.translate = FALSE ) +                
scale_alpha_discrete(guide = "none") + theme(
    plot.title = element_text(size = 16),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )
    ggsave(filename = paste0("/Users/debanjansaha/Desktop/pc_plots_v4/", pid, ".pdf" ), plot = p)
  return(p)
})

 
 

```


```{r}

#Use Hierarchical cluster tree from CopyKat - not ultrametric so allows distance metric
tree <- readRDS(file = "/Users/debanjansaha/Desktop/mCRPC_analysis/R/crpc_analysis/results/epi20000tme20000seed123_copykat_clustering_results.rds")


#Helper to compute normalized shannon entropy 
compute_pacc_entropy <- function(groups, labels) {
  # labels: named vector (tip -> PACCS status)
  # groups: named vector (tip -> CN cluster)

  pacc_tips <- names(labels)[labels == "PACCs"]
  if (length(pacc_tips) < 2) return(NA_real_)

  #Get groups of each tips from clustering
  clades <- groups[pacc_tips]
  #Proportion calculation of each group
  p <- table(clades)
  p <- p / sum(p)
  
  #Information entropy
  H <- -sum(p * log(p))
  #Normalization to max entropy
  H_norm <- H / log(length(p))

  return(H_norm)
}

patients <- unique(pacc_meta$patient)

entropy_results <- lapply(patients, function(pid) {

  cat("Patient:", pid, "\n")

  ## ---- subset cells ----
  ids <- pacc_meta %>%
    filter(patient == pid, cell_name %in% medoid_cells) %>%
    pull(cell_name) %>%
    gsub("-", "\\.", .)

  m <- mediods[, colnames(mediods) %in% ids, drop = FALSE]

  if (ncol(m) < 5) {
    message("Skipping patient ", pid)
    return(NULL)
  }

  #Get distance, in this case use manhattan L1  
  d <- dist(t(m), method = "manhattan")

  #Hclustering
  hc <- hclust(d, method = "average") 

 #Cut the tree to get groupings
  groups <- cutree(hc, h = quantile(hc$height, 0.6))

  #Label PCs
  labels <- pacc_meta %>%
    filter(patient == pid) %>%
    mutate(label = gsub("-", "\\.", cell_name)) %>%
    filter(label %in% names(groups)) %>%
    dplyr::select(label, PACCS) %>%
    deframe()

  #Get normalized entropy per patient generated tree and return 
  entropy <- compute_pacc_entropy(groups, labels)

  return(data.frame(
    patient = pid,
    entropy = entropy,
    n_cells = length(groups),
    n_paccs = sum(labels == "PACCs", na.rm = TRUE)
  ))
})

entropy_df <- bind_rows(entropy_results)

pdf("/Users/debanjansaha/Desktop/pc_plots_v4/entropy_paccs.pdf", width = 10)
ggplot(entropy_df,aes( y = patient, x = entropy, size = n_paccs, color = n_cells)) + geom_point() + theme_bw(base_size = 20) + scale_color_viridis_c(option = "D", name = "Number of Cells") + xlab("Normalized Shannon Entropy") + ylab("Patient ID") + scale_size_continuous(
    name = "Number of PCs",          
    range = c(2, 8),                
    breaks = c(0, 50, 100)         
  )
dev.off()
```




#Plotting of training and validation results
```{r}
#Plot AUC of training data (PCA with C)
library(pROC)
pc3_c_train_auc <- read.table(file = "/Users/debanjansaha/Desktop/pc_plots_v4/paccs_plots_v2/pca_c_class_train_results.tsv", header = T)

p1 <- pc3_c_train_auc %>% mutate(Fold = 1:n()) %>%
  pivot_longer(
    cols = -Fold,
    names_to = "Model",
    values_to = "AUC", 
  ) %>% ggplot(aes(x = Fold, y = AUC, fill = Model)) + scale_fill_viridis_d(option = "B") + geom_col(position = "dodge") + theme_bw(base_size = 20) + ylim(c(0,1))

#Plot Feature importances in training data
pc3_c_feature <- read.table(file = "/Users/debanjansaha/Desktop/pc_plots_v4/paccs_plots_v2/pca_c_feature_import.tsv", header = T)

p2 <- pc3_c_feature %>% ggplot(aes(x = Feature, y = Importance, fill = Model)) + scale_fill_viridis_d(option = "E") + geom_col(position = "dodge") + theme_bw(base_size = 20) + ylim(c(0,1))  + scale_x_discrete(labels = c(
    cnv_total    = "Total CNV",
    G2M.Score    = "G2M Score",
    nCount_RNA   = "UMI Count",
    nFeature_RNA = "Gene Count",
    Percent.MT   = "% Mito",
    S.Score      = "S Phase Score"
  ))

#Plot AUC of validation data (PCA with D, MDA with C and D)
pc3_d_valid_auc <- read.table(file = "/Users/debanjansaha/Desktop/pc_plots_v4/paccs_plots_v2/pca_d_results.tsv", header = T)
mda_c_valid_auc <- read.table(file = "/Users/debanjansaha/Desktop/pc_plots_v4/paccs_plots_v2/mda_c_results.tsv", header = T)
mda_d_valid_auc <- read.table(file = "/Users/debanjansaha/Desktop/pc_plots_v4/paccs_plots_v2/mda_d_results.tsv", header = T)



p3 <- rbind(pc3_d_valid_auc %>%
  dplyr::select(ends_with("WGD_prob")) %>%
  map_dbl(~ as.numeric(auc(pc3_d_valid_auc$label, .x)))  %>%  
    enframe(name = "model", value = "AUC") %>% mutate(Data = "PC3 Docetaxel"), 
  mda_c_valid_auc %>% 
    dplyr::select(ends_with("WGD_prob")) %>%
  map_dbl(~ as.numeric(auc(mda_c_valid_auc$label, .x))) %>%
  enframe(name = "model", value = "AUC")  %>% mutate(Data = "MDA Cisplatin") )  %>% 
  rbind(
    mda_d_valid_auc %>% dplyr::select(ends_with("WGD_prob")) %>%
  map_dbl(~ as.numeric(auc(mda_d_valid_auc$label, .x))) %>%
  enframe(name = "model", value = "AUC")   %>% mutate(Data = "MDA Docetaxel")
  ) %>% ggplot(aes(y = model, x = AUC, fill = Data)) + scale_fill_viridis_d(end = 0.3) + geom_col(position = "dodge") + theme_bw(base_size = 20) + xlim(c(0,1))   + scale_y_discrete(labels = c(
    XGBoost_WGD_prob    = "XGBoost",
    SVM_WGD_prob    = "SVM",
    RandomForest_WGD_prob   = "RandomForest",
    NeuralNet_WGD_prob = "NeuralNet",
    LogisticRegression_WGD_prob   = "LogisticRegression"
  ))
 
p4 <- rbind(
  dplyr::select(pc3_d_valid_auc, label, ends_with("WGD_prob")) %>% 
    mutate(Data = "PC3 Docetaxel"),
  dplyr::select(mda_c_valid_auc, label, ends_with("WGD_prob")) %>% 
    mutate(Data = "MDA Cisplatin"),
  dplyr::select(mda_d_valid_auc, label, ends_with("WGD_prob")) %>% 
    mutate(Data = "MDA Docetaxel")
)  %>% 
  pivot_longer(cols = -c(label, Data),
               names_to = "Model",
               values_to = "PACCs_Probability") %>% 
  mutate(Model = recode(Model,
  "XGBoost_WGD_prob"    = "XGBoost",
    "SVM_WGD_prob"    = "SVM",
    "RandomForest_WGD_prob"   = "RandomForest",
    "NeuralNet_WGD_prob" = "NeuralNet",
    "LogisticRegression_WGD_prob"   = "LogisticRegression"
)) %>%
  ggplot(aes(x = as.factor(label),
             y = PACCs_Probability,
             fill = as.factor(label))) +
  geom_violin(scale = "width", trim = FALSE) +
  facet_grid(Data ~ Model) +
  theme_bw(base_size = 20) +
  scale_fill_viridis_d(option = "E", end = 0.5, name = "Cell Type",
    labels = c("0" = "Non-PCs",
               "1" = "PCs")) + xlab("Cell Type") + ylab("PCs probability")  + theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )

ggsave(filename = "/Users/debanjansaha/Desktop/pc_plots_v4/train_AUC.pdf", plot = p1, width = 10)
ggsave(filename = "/Users/debanjansaha/Desktop/pc_plots_v4/train_import.pdf", plot = p2, width = 10)
ggsave(filename = "/Users/debanjansaha/Desktop/pc_plots_v4/valid_AUC.pdf", plot = p3, width = 10)
ggsave(filename = "/Users/debanjansaha/Desktop/pc_plots_v4/valid_prob.pdf", plot = p4, width = 15)


```

